Reduced MHD Notes

### 2025-01-27 Notes/Notes.tex
Beta Boundary Conditions

\begin{lemma} \label{Homo Harmonic}
    The unique harmonic function with h BCs is $f=0$. 
\end{lemma}
\begin{proof}
    General form for harmonic function on disc is 
    \begin{align}
        f(r,\theta) &= \frac{A_0}{2} + \sum_{n=1}^\infty \pth{ A_nr^n\cos n\theta + B_nr^n\sin n\theta}.
    \end{align}
    If $f$ has hD BCs, $A_n=B_n=0$ to eliminate dependence on $\theta$, and $A_0=0$ to set function value to zero. If $f$ has hN BCs, then 
    \begin{align}
        \bm{n}\cdot\np f(r,\theta) = \deriv{f}{r}(r,\theta) &= \sum_{n=1}^\infty \pth{ nA_nr^{n-1} \cos n\theta + nB_nr^{n-1}\sin n\theta} = 0
    \end{align}
    requires $A_n=B_n=0$. Then, $\iint_D f\ dxdy = \frac{A_0}{2}xy = 0$ requires $A_0=0$. Thus $f(r,\theta)=0$ everywhere. 
\end{proof}
\begin{proof}
    MUCH easier to use principle that harmonic functions take max/min on boundary. If boundary is 0, then $f=0$. If deriv on boundary is 0, then the function must be constant (or else the function would achieve a max or min somewhere else than the boundary). Only constant with average value zero is $f=0$. 
\end{proof}

@@

\begin{definition}
    Define $H^D_g$ and $H^N_g$ as the unique harmonic functions with the same Dirichlet and Neumann boundary data as $g$ respectively. That is, $\lap H^D_g \lap H^N_g = 0$,  
    \begin{align}
        H^D_g = g \quad\text{and}\quad \bm{n}\cdot\np H^N_g = \bm{n}\cdot\np g \quad\text{on}\quad \p D, \quad\text{and}\quad &&  \iint_D H^N_g\ dxdy = \iint_D g\ dxdy. 
    \end{align}
\end{definition}
Existence of $H_g$ requires that $\int_{\p D} \bm{n}\cdot\np H_g\ dl = \int_{\p D} \bm{n}\cdot\np g\ dl = 0$ by (Lemma \ref{Neumann Problem}). This can be rewritten as $\int_{\p D} \bm{n}\cdot\np g\ dl = \iint_D s\ dxdy = 0$ for $s=\lap g$. 
By (Lemma \ref{Unique Laplace}), $H_g$ is unique becuase it is harmonic with boundary conditions. 
Because $H_{ag_1+bg_2}$ is unique, and because $\lap(aH_{g_1}+ bH_{g_2}) = a\lap H_{g_1} + b\lap H_{g_2} = 0$ (assuming $H_{g_1}$ and $H_{g_2}$ both exist for a given $g_1$ and $g_2$), $H_{ag_1+bg_2} = aH_{g_1} + bH_{g_2}$ is linear.

@@

\subsection{Boundary Conditions for $\beta_\para$}
Divergence theorem argument: Consider a flux tube going from $z_1$ to $z_2$. Because $\nabla\cdot\bm{B}=0$,  
$$\iiint_V \pth{\nabla\cdot \bm{B}} dV = \iint_{\p V} \pth{\bm{B}\cdot \bm{n}} dA = 0.$$
However, we specifically have the boundary condition that $\bm{B}\cdot\bm{n}=0$ on $\p Q$, so we only need to consider contributions from the poloidal discs. Then the resulting integral becomes 
$$\iint_D B_z(x,y,z_2)dxdy = \iint_D B_z(x,y,z_1) dxdy.$$
Thus, we are free to set $B_z(x,y,z) = B_z(x,y) = B_0 \pth{1 + \ep\beta_\para(x,y)}$, where only $B_0$ contributes to the total flux. That is,  
$$\iint_D B_zdxdy = B_0A + \ep B_0\iint_D \beta_\para(x,y)dxdy = B_0A.$$
This shows we can choose $\beta_\para(x,y)$ to have boundary conditions where $\iint_D \beta_\para\ dxdy = 0$.

@@

Assume solutions are all smooth $C^\infty$ in time/``regular." Then eg. deriv of $v\cdot n=v_\perp\cdot n=0$ is $\p_t v\cdot n=0$ which is technically another boundary condition (we have an expression for $\p_t v=$, but no reason to believe a priori that it is $=0$, so this is assumed). This is required when I try to pull $\lap$ off of $\lap\p_t\phi = \np\cdot\p_tv_\perp$. Lemma 1 says existence of solutions to this Poisson equation $\lap f=s$ require $\iint_D s\ dxdy=0 = \oint_{\p D} n\cdot\np f\ dl =0$. In this case, solutions require $\iint_D \np\cdot\p_t v_\perp\ dxdy = \oint_{\p D} (\p_tv_\perp \cdot n)\ dl =0$, which holds if $v\in C^\infty$ so that we can take $\p_t (v_\perp \cdot n) = 0$.

>>>>

\begin{lemma} \label{Homo Harmonic}
    The unique harmonic function on the disc with h BCs is $f=0$. 
\end{lemma}
\begin{proof}
    General form for harmonic function on disc is 
    \begin{align}
        f(r,\theta) &= \frac{A_0}{2} + \sum_{n=1}^\infty \pth{ A_nr^n\cos n\theta + B_nr^n\sin n\theta}.
    \end{align}
    If $f$ has hD BCs, $A_n=B_n=0$ to eliminate dependence on $\theta$, and $A_0=0$ to set function value to zero. If $f$ has hN BCs, then 
    \begin{align}
        \bm{n}\cdot\np f(r,\theta) = \deriv{f}{r}(r,\theta) &= \sum_{n=1}^\infty \pth{ nA_nr^{n-1} \cos n\theta + nB_nr^{n-1}\sin n\theta} = 0
    \end{align}
    requires $A_n=B_n=0$. Then, $\iint_D f\ dxdy = \frac{A_0}{2}xy = 0$ requires $A_0=0$. Thus $f(r,\theta)=0$ everywhere. 
\end{proof}
\begin{proof}
    MUCH easier to use principle that harmonic functions take max/min on boundary. If boundary is 0, then $f=0$. If deriv on boundary is 0, then the function must be constant (or else the function would achieve a max or min somewhere else than the boundary). Only constant with average value zero is $f=0$. 
\end{proof}

@@

\begin{definition}
    Define $H^D_g$ and $H^N_g$ as the unique harmonic functions with the same Dirichlet and Neumann boundary data as $g$ respectively. That is, $\lap H^D_g = \lap H^N_g = 0$,  
    \begin{align}
        H^D_g = g \quad\text{and}\quad \bm{n}\cdot\np H^N_g = \bm{n}\cdot\np g \quad\text{on}\quad \p D, \quad\text{and}\quad &&  \iint_D H^N_g\ dxdy = \iint_D g\ dxdy. 
    \end{align}
\end{definition}
Existence of $H_g$ requires that $\int_{\p D} \bm{n}\cdot\np H_g\ dl = \int_{\p D} \bm{n}\cdot\np g\ dl = 0$ by (Lemma \ref{Neumann Problem}). This can be rewritten as $\int_{\p D} \bm{n}\cdot\np g\ dl = \iint_D s\ dxdy = 0$ for $s=\lap g$. 
By (Lemma \ref{Unique Laplace}), $H_g$ is unique becuase it is harmonic with boundary conditions. 
Because $H_{ag_1+bg_2}$ is unique, and because $\lap(aH_{g_1}+ bH_{g_2}) = a\lap H_{g_1} + b\lap H_{g_2} = 0$ (assuming $H_{g_1}$ and $H_{g_2}$ both exist for a given $g_1$ and $g_2$), $H_{ag_1+bg_2} = aH_{g_1} + bH_{g_2}$ is linear.

@@

\subsection{Compatibility Conditions/Regularity}
We have been using $\GN:L^2(D)\to\mathcal{N}(D)$ to find $f$ from $s = \lap f$ with the requirement that $\iint_D s\ dA = 0$ from Lemma \ref{Neumann Problem}. So really, we mean $\GN:\mathcal{F}\to\mathcal{N}$, where $s\in\mathcal{F}\subset L^2(D)$ if and only if $\iint_D s\ dA = \int_{\p D} \np f\cdot\bm{n}\ dl = 0$.

For $\lap\phi = \np\cdot\bm{\nu}_\perp$, this works because we have the boundary condition $\bm{v}\cdot\bm{n} = 0$ on $\p Q$, or $\bm{v}_\perp\cdot\bm{n}_\perp = 0$ on $\p D$: $\iint_D \np\cdot\bm{\nu}_\perp\ dA = \int_{\p D} \bm{\nu}_\perp\cdot\bm{n}_\perp\ dl = 0$, so $s\in\mathcal{F}$.

When inverting $\lap\p_\tau\phi = \np\cdot\p_\tau\bm{\nu}_\perp$, we instead require $\iint_Q \np\cdot\p_\tau\bm{\nu}_\perp\,dA = \int_{\p Q} \p_\tau\bm{\nu}_\perp\cdot\bm{n}_\perp\,dl = 0$. If $\bm{\nu}_\perp(t)\in C^\infty$ varies smoothly with time, we may take $\p_\tau\pth{\bm{\nu_\perp}\cdot\bm{n}_\perp} = \p_\tau\bm{\nu}_\perp\cdot\bm{n}_\perp = 0$. If $\bm{\nu}_\perp$ does not have some notion of ``regularity'' in time however, this is not necessarily the case, and $\np\cdot\p_\tau\bm{\nu}_\perp$ may not $\in\mathcal{F}$. This distinction becomes clear when we consider that our expression for $\p_\tau\bm{\nu}_\perp$ in \eqref{nuperpevolution}; setting $\p_\tau\bm{\nu}_\perp\cdot\bm{n}_\perp = 0$ seems to imply additional conditions on $\beta_\para$, $\bm{\beta}_\perp$, $\bm{\nu}_\perp$, and $\pi$. Thus $\p_\tau s\in\mathcal{F}$, only follows from $s\in\mathcal{F}$ if $s$ obeys the above ``compatibility condition.''

\subsection{Boundary Conditions for $\beta_\para$}
Divergence theorem argument: Consider a flux tube $T\subset Q$ going from $Z_1$ to $Z_2$. Because $\nabla\cdot\bm{B}=0$,  
\begin{align*}
\iiint_T \pth{\nabla\cdot \bm{B}} dV = \iint_{\p T} \pth{\bm{B}\cdot \bm{n}} dA = 0.
\end{align*}
However, we specifically have the boundary condition that $\bm{B}\cdot\bm{n}=0$ on $\p Q$, so we only need to consider contributions from the poloidal discs. Then the resulting integral becomes 
\begin{align*}
\iint_D B_z\pth{X,Y,Z_2}\,dA = \iint_D B_z\pth{X,Y,Z_1}\,dA
\end{align*}
for arbitrary $Z_1$ and $Z_2$, which implies $\iint_D \p_ZB_z\,dA = 0$. When we set $B_z = B_0\pth{1+ \ep \beta_\para}$, this condition becomes $\iint_D \p_Z\beta_\para\,dA=0$. This is necessary when we try to invert $\lap\Phi = -\ep\p_Z\beta_\para$, which requires $\p_Z\beta_\para\in\mathcal{F}$.

$\beta_\para$ has a poloidal average, $\bar{\beta}_\para = \frac{1}{A}\iint_D \beta_\para\,dA$.
This lets us decompose $\beta_\para\pth{X,Y,Z} = \bar{\beta}_\para\pth{Z} + \til{\beta}_\para\pth{X,Y,Z}$, where $\til{\beta}_\para\in\mathcal{F}$ obeys $\iint_D \til{\beta}_\para\,dA = 0$ (this is because $= \iint_D \beta_\para\,dA - \iint_D \bar{\beta}_\para\pth{Z}\,dA = \bar{\beta}_\para A - \bar{\beta}_\para A = 0$). Then, the condition above becomes $\p_Z\bar{\beta}_\para = \frac{1}{A}\iint_D \p_Z\beta_\para\,dA = 0$, implying that $\bar{\beta}_\para\pth{Z}=$ constant, so $\iint_D \p_Z\beta_\para\,dA = \iint_D \p_Z\til{\beta}_\para\,dA=0$.

@@
